---
title: "Milestone report -  Capstone Project"
subtitle: "Coursera Data Science Specialization"
author: "Filipe Pais Lenfers"
date: "14 de março de 2015"
output: html_document
---

# Synopsis

The Capstone Project objetive is to build a predictive text model from [HC Corpora](http://www.corpora.heliohost.org) database. The model should suggest the next word based on the previous words used. The text documents are provided from diferent three web sources: blogs, twitter and news articles. This report demonstrates the preliminary exploration of the data and the possible ways to build the prediction algorithm.

# Exploratory Analisys

## Obtaining the data

The data has been download from this link: [https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip](https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip), and uncompressed. 

```{r, eval=FALSE, echo=FALSE}
destination.file <- "Coursera-SwiftKey.zip"
link <- "http://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip"

download.file(link, destination.file) #Download the data
unzip(destination.file) # Uncompress the data
file.rename("final","data") # Rename the final directory to data
```

There are four languages in this data set. On on each directory.
```{r}
list.files("data")
```

We will concetrate our analysis on the english data, there are 3 files from diferent sources in data:
```{r}
list.files("data/en_US")
```

The source of the files are from:

* Blogs
* Newspappers
* Twitter

More details can be found in [HC Corpora](http://www.corpora.heliohost.org/aboutcorpus.html) website.

## Data Cleanning and Preprocessing

We will concentrate our analysis in the english files.
```{r,cache=TRUE}
blogs <- readLines("data/en_US/en_US.blogs.txt", encoding="UTF-8")
twitter <- readLines("data/en_US/en_US.twitter.txt", encoding="UTF-8")
news  <- readLines("data/en_US/en_US.news.txt", encoding="UTF-8")
```


## Basic Data Analysis

