---
title: "Milestone report -  Capstone Project"
subtitle: "Coursera Data Science Specialization"
author: "Filipe Pais Lenfers"
date: "14 de março de 2015"
output: html_document
---

# Synopsis

The Capstone Project objetive is to build a predictive text model from [HC Corpora](http://www.corpora.heliohost.org) database. The model should suggest the next word based on the previous words used. The text documents are provided from diferent three web sources: blogs, twitter and news articles. This report demonstrates the preliminary exploration of the data and the possible ways to build the prediction algorithm.

# Exploratory Analisys

## Libraries

Libraries used in this study.
```{r}
library(stringi)
library(tm)
library(RWeka)
```


## Obtaining the data

The data has been download from this link: [https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip](https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip), and uncompressed. 

```{r, eval=FALSE}
destination.file <- "Coursera-SwiftKey.zip"
link <- "http://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip"

download.file(link, destination.file) #Download the data
unzip(destination.file) # Uncompress the data
file.rename("final","data") # Rename the final directory to data
```

There are four languages in this data set. On on each directory.
```{r}
list.files("data")
```

We will concetrate our analysis on the english data, there are 3 files from diferent sources in data:
```{r}
list.files("data/en_US")
```

The source of the files are from:

* Blogs
* Newspappers
* Twitter

More details can be found in [HC Corpora](http://www.corpora.heliohost.org/aboutcorpus.html) website.

# Files Analysis

Lets load the three files to obtain some basic information about then.
```{r,cache=TRUE,warning=FALSE}
blog.file <- "data/en_US/en_US.blogs.txt"
twitter.file <- "data/en_US/en_US.twitter.txt"
news.file <- "data/en_US/en_US.news.txt"
blogs <- readLines(blog.file, encoding="UTF-8")
twitter <- readLines(twitter.file, encoding="UTF-8")
news  <- readLines(news.file, encoding="UTF-8")
```

Blog file:

* File name: `r blog.file`.
* File size: `r round(file.info(blog.file)$size/1024/1024,2)` MB.
* Lines and chars statistics on the file:
```{r, cache=TRUE}
data.frame(t(stri_stats_general(blogs)))
```
* Summary of the char quantities per line:
```{r,cache=TRUE}
summary(sapply(blogs,FUN=nchar))
```
* Words summary:
```{r,cache=TRUE}
summary(stri_count_words(blogs))
```

Twitter file:

* File name: `r twitter.file`.
* File size: `r round(file.info(twitter.file)$size/1024/1024,2)` MB.
```{r, cache=TRUE}
data.frame(t(stri_stats_general(twitter)))
```
* Summary of the char quantities per line:
```{r,cache=TRUE}
summary(sapply(twitter,FUN=nchar))
```
* Words summary:
```{r,cache=TRUE}
summary(stri_count_words(twitter))
```

News file:

* File name: `r news.file`.
* File size: `r round(file.info(news.file)$size/1024/1024,2)` MB.
```{r, cache=TRUE}
data.frame(t(stri_stats_general(news)))
```
* Summary of the char quantities per line:
```{r,cache=TRUE}
summary(sapply(news,FUN=nchar))
```
* Words summary:
```{r,cache=TRUE}
summary(stri_count_words(news))
```

## Data Cleanning and Preprocessing

As we could see in the previous section the files are too big, so we will extract some samples to work with.
```{r}
sample.size <- 100000
```

```{r, eval=FALSE}
sample.blogs <- sample(blogs,sample.size)
sample.twitter <- sample(twitter,sample.size)
sample.news <- sample(news,sample.size)
```

* sample.blogs has `r round(sample.size/length(blogs)*100.0,2)`% of the original data.
* sample.twitter has `r round(sample.size/length(twitter)*100.0,2)`% of the original data.
* sample.news has `r round(sample.size/length(news)*100.0,2)`% of the original data.

To garantie the reproducibility of this research we will save the samples and load then as needed.

```{r,eval=FALSE}
save(sample.blogs,sample.twitter,sample.news,file="samples.RData")
```

```{r}
load("samples.RData")
```

Lets convert the encoding of the characters, removing the non-convertible characters.
```{r, cache=TRUE}
sample_blogs <- iconv(sample_blogs, "latin1", "ASCII", sub="")
sample_twitter <- iconv(sample_twitter, "latin1", "ASCII", sub="")
sample_news <- iconv(sample_twitter, "latin1", "ASCII", sub="")
```

Now we need to clean the text, making all text lower case, removing pontuation, removing the number, removing the [stop words](http://en.wikipedia.org/wiki/Stop_words), removing profanities and striping the white spaces.We also (stem)[http://en.wikipedia.org/wiki/Stemming] the text.
The list of profanity words was obtained from [https://github.com/shutterstock/List-of-Dirty-Naughty-Obscene-and-Otherwise-Bad-Words/blob/master/en](https://github.com/shutterstock/List-of-Dirty-Naughty-Obscene-and-Otherwise-Bad-Words/blob/master/en).
This steps will make the data easier to analize.
```{r, cache=TRUE}
profanity.words <- readLines("en_profanity_words.txt")

corpus.blogs <- Corpus(VectorSource(list(sample.blogs)))
corpus.blogs <- tm_map(corpus.blogs, content_transformer(tolower))
corpus.blogs <- tm_map(corpus.blogs, content_transformer(removePunctuation))
corpus.blogs <- tm_map(corpus.blogs, content_transformer(removeNumbers))
corpus.blogs <- tm_map(corpus.blogs, removeWords, stopwords("english"))
corpus.blogs <- tm_map(corpus.blogs, removeWords, profanity.words)
corpus <- tm_map(corpus.blogs, stripWhitespace)

corpus.blogs <- tm_map(corpus.blogs, stemDocument, language='english')
```

```{r, cache=TRUE}
corpus.twitter <- Corpus(VectorSource(list(sample.twitter)))
corpus.twitter <- tm_map(corpus.twitter, content_transformer(tolower))
corpus.twitter <- tm_map(corpus.twitter, content_transformer(removePunctuation))
corpus.twitter <- tm_map(corpus.twitter, content_transformer(removeNumbers))
corpus.twitter <- tm_map(corpus.twitter, removeWords, stopwords("english"))
corpus.twitter <- tm_map(corpus.twitter, removeWords, profanity.words)
corpus.twitter <- tm_map(corpus.twitter, stripWhitespace)

corpus.twitter <- tm_map(corpus.twitter, stemDocument, language='english')
```

```{r, cache=TRUE}
corpus.news <- Corpus(VectorSource(list(sample.news)))
corpus.news <- tm_map(corpus.news, content_transformer(tolower))
corpus.news <- tm_map(corpus.news, content_transformer(removePunctuation))
corpus.news <- tm_map(corpus.news, content_transformer(removeNumbers))
corpus.news <- tm_map(corpus.news, removeWords, stopwords("english"))
corpus.news <- tm_map(corpus.news, removeWords, profanity.words)
corpus.news <- tm_map(corpus.news, stripWhitespace)

corpus.news <- tm_map(corpus.news, stemDocument, language='english')
```

The data is prepared to be analized.

## Data Analysis

### Unigram analysis

### Bigram analysis
